---
title: "Reproducible Research: Peer Assessment 1"
author: "Turing"
date: "2022-12-29"
output: html_document
---

## Loading and preprocessing the data
  
  First, let's unzip the data file and read the data into R.
    
```{r}
outdir <- getwd()
unzip("activity.zip", exdir = outdir)
mydata <- read.csv("activity.csv", header = TRUE)
```
    
  The structure, first 6 lines, and class of the data are shown as the follows:
    
```{r}
str(mydata)
head(mydata)
class(mydata)
```

## What is mean total number of steps taken per day?
  
  We grouped the data by date and summarized the total number of steps taken each day.

```{r}
library(dplyr)
library(lubridate)
q1data <- mydata %>%
    mutate(date = ymd(date)) %>%
    group_by(date) %>%
    summarise(steps = sum(steps, na.rm = TRUE))
head(q1data)
```

  The histogram of the total number of steps taken each day is shown as below:

```{r}
hist(q1data$steps, breaks = 20, 
     xlab = "Steps", ylab = "Number of days",
     main = "The total number of steps taken each day")
```

  The mean number of steps taken per day is:

```{r}
mean(q1data$steps, na.rm = TRUE)
```

  The median number of steps taken per day is:

```{r}
median(q1data$steps, na.rm = TRUE)
```

## What is the average daily activity pattern?

  We grouped the data by intervals and summarized the total number of steps taken during each interval.
  
```{r}
q2data <- mydata %>%
    group_by(interval) %>%
    summarise(steps = mean(steps, na.rm = TRUE))
head(q2data)
```

  The time series plot of the average number of steps taken during each intervals is shown as below:

```{r}
within(q2data, plot(interval, steps, type = "l",
                    ylab = "Average number of steps across all days",
                    main = "The average daily activity pattern"))
```

  The 5-minute interval that, on average, contains the maximum number of steps is:

```{r}
q2data$interval[which.max(q2data$steps)]
```

## Imputing missing values
  
  The total number of missing values in the dataset is:

```{r}
totalMV <- sum(is.na(mydata$steps))
totalMV
```

  To impute missing values in the dataset, I write a functon that takes the dataset with missing values and the dataset which contains the total number of steps taken per day as parameters. Specifically, I use the mean number of steps for that day to fill in the missing values on that day.

```{r}
impdata <- mydata %>%
    mutate(date = ymd(date)) %>%
    group_by(date) %>%
    summarise(steps = mean(steps, na.rm = TRUE))

imputeMV <- function(ori_data, imp_data) {
    MVid <- which(is.na(ori_data$steps)==1)
    ori_data <- mutate(ori_data, date = ymd(date))
    for (i in 1:length(MVid)) {
        fillid <- which(imp_data$date==ori_data$date[MVid[i]])
        ori_data$steps[MVid[i]] <- imp_data$steps[fillid]
    }
    return(ori_data)
}

mydata2 <- imputeMV(mydata, impdata)
```

  The histogram of the total number of steps taken each day after missing values are imputed is as below

```{r}
q3data <- mydata2 %>%
    group_by(date) %>%
    summarise(steps = sum(steps, na.rm = TRUE))
hist(q3data$steps, breaks = 20, 
     xlab = "Steps", ylab = "Number of days",
     main = "The total number of steps taken each day (filled data)")
```

  The mean and median number of steps taken per day are:
  
```{r}
mean(q3data$steps, na.rm = TRUE)
median(q3data$steps, na.rm = TRUE)
```
  
  As we can see, the mean and median number of steps taken per day in the dataset after imputing missing values are the same as they are in the original dataset.
  
## Are there differences in activity patterns between weekdays and weekends?

  Create a new factor variable in the dataset with two levels – “weekday” and “weekend” indicating whether a given date is a weekday or weekend day.
  
```{r}
Sys.setlocale("LC_TIME", "English")
mydata2$weekdays <- weekdays(mydata2$date)
mydata2$weekdays[which(mydata2$weekdays%in%c("Monday","Tuesday","Wednesday","Thursday","Friday" ))] <- "weekday"
mydata2$weekdays[which(mydata2$weekdays%in%c("Saturday","Sunday"))] <- "weekend"
mydata2$weekdays <- as.factor(mydata2$weekdays)
head(mydata2)
```

  The time series plot of the 5-minute interval and the average number of steps taken, averaged across all weekday days or weekend days is shown as below:

```{r}
q4data <- mydata2 %>%
    group_by(interval,weekdays) %>%
    summarise(steps = mean(steps, na.rm = TRUE))
library(lattice)
xyplot(steps~interval|weekdays, q4data, layout = c(1,2), type = "l")
```